Simulating High-Dimensional Multivariate Data

### 2021-03-01 010-background.Rmd
added GSDist ref. Did built-in spell check

We illustrate our methodology using a well-studied Breast Invasive Carcinoma (BRCA) data set housed in The Cancer Genome Atlas (TCGA; see Acknowledgments).
For simplicity, we only consider high expressing genes. In turn, we begin by filtering to retain the top `r d` of the highest-expressing genes (in terms of median expression) of the oover 20,000 gene measurements from $N=`r nrow(example_brca)`$ patients' tumor samples. This gives a great number of pairwise dependencies among the marginals (specifically, $`r choose(d,2)`$ correlation parameters). Table \@ref(tab:ch010-realDataTab) displays RNA-seq counts for three selected high-expressing genes for the first five patients' breast tumor samples.

>>>>

We illustrate our methodology using a well-studied Breast Invasive Carcinoma (BRCA) data set housed in The Cancer Genome Atlas (TCGA; see Acknowledgments).
For simplicity, we only consider high expressing genes. In turn, we begin by filtering to retain the top `r d` of the highest-expressing genes (in terms of median expression) of the over 20,000 gene measurements from $N=`r nrow(example_brca)`$ patients' tumor samples. This gives a great number of pairwise dependencies among the marginals (specifically, $`r choose(d,2)`$ correlation parameters). Table \@ref(tab:ch010-realDataTab) displays RNA-seq counts for three selected high-expressing genes for the first five patients' breast tumor samples.

### 2021-03-01 020-simulation-algorithm.Rmd
added GSDist ref. Did built-in spell check

*Step 1*. The first two descriptions of the *Mapping step* employ the closed-form relationships between $\rho_S$ and $\tau$ with $\rho_P$ for bivariate normal random variables via Equations \@ref(eq:convertKendall) and \@ref(eq:convertSpearman), respectively (implemented as `cor_covert`). Initializing our algorthim to match the nonparametric correlations by computing these equations for all pairs is computationally trivial. For the computationally expensive process to match Pearson correlations, we approximate Equation \@ref(eq:pearsonIntegralRelation) for all pairs of margins. To this end, we implement the approximation scheme introduced by @xiao2019matching. Briefly, the many double integrals of the form in \@ref(eq:pearsonIntegralRelation) are approimated by weighted sums of Hermite polynomials.

>>>>

*Step 1*. The first two descriptions of the *Mapping step* employ the closed-form relationships between $\rho_S$ and $\tau$ with $\rho_P$ for bivariate normal random variables via Equations \@ref(eq:convertKendall) and \@ref(eq:convertSpearman), respectively (implemented as `cor_covert`). Initializing our algorithm to match the nonparametric correlations by computing these equations for all pairs is computationally trivial.

For the computationally expensive process to match Pearson correlations, we approximate Equation \@ref(eq:pearsonIntegralRelation) for all pairs of margins. To this end, we implement the approximation scheme introduced by @xiao2019matching. Briefly, the many double integrals of the form in \@ref(eq:pearsonIntegralRelation) are approximated by weighted sums of Hermite polynomials. Matching coefficients for pairs of continuous distributions is made tractable by this method, but for discrete distributions (especially discrete distributions with large support sets or infinite support), the problem is not so simple or efficient.

Our solution is to approximate discrete distributions by a continuous distribution. However the question is then what distribution to use. We found that Generalized S-Distributions [@muino2006gs] solve this problem by approximating a wide range of unimodal distributions, both continuous and discrete. Using the GS-Distribution approximation of discrete distributions in Pearson matching scheme above yields favorable results.

### 2021-03-01 080-appendix.Rmd
added GSDist ref. Did built-in spell check

\noindent Consider a bivariate example where we have a correlated $(Y_1, Y_2)^\top$ with $Y_1\sim N(\mu_1, \sigma_1^2)$ (normal) and $Y_2\sim LN(\mu_2, \sigma_2^2)$ (lognormal). First, let us note that when we get such a vector from BigSimR package then in fact it can be represented as

@@

Plugging in the estimands of these quantities from Section \@ref(package), we see that

>>>>

\noindent Consider a bivariate example where we have a correlated $(Y_1, Y_2)^\top$ with $Y_1\sim N(\mu_1, \sigma_1^2)$ (normal) and $Y_2\sim LN(\mu_2, \sigma_2^2)$ (lognormal). First, let us note that when we get such a vector from bigsimr package then in fact it can be represented as

@@

Plugging in the estimates of these quantities from Section \@ref(package), we see that
